HCM SCD risk algorithm cost-effectiveness analysis

### 2021-04-20 HCM.SCD.CEanalysis/inst/rmarkdown/draft.Rmd
replace constants cost inputs with functions to sample costs at runtime

---
title: "HCM SCD risk algorithm cost-effectiveness analysis"
author: "Nathan Green"
date: "09/11/2020"
output: pdf_document
editor_options: 
  markdown: 
    wrap: sentence
references:
- id: mahony2014
  title: A novel clinical risk prediction model for sudden cardiac death in hypertrophic cardiomyopathy (HCM risk-SCD)
  author:
  - family: O'Mahony
    given: C
  container-title: Eur Heart
  volume: 35
  URL: 'https://pubmed.ncbi.nlm.nih.gov/24126876/'
  DOI:
  issue: 30
  page: 2010-20
  type: article-journal
  issued:
    year: 2014
---

@@

Risk score algorithm details see [@mahony2014].

@@

 Name | Value | Parameter | Source 
------|-------|-----------|--------
| _Health_ ||||
| Manage with ICD | 0.637 QALY/year | $h_1$ | Noyes 2007 |
| Shock | ? utility losss | $h_2$ |  |
| HCM | ? QALY/year | $h_3$ |  |
| Death | 0 QALY/year | $h_4$, $h_5$ |  |
|       |      |      |     |
| _Cost_ ||||
| Manage with ICD | ? | $c_1$ |  |
| Non-fatal HCM event | £22,880 | $c_2$ | UK Stroke Assoc. |
| HCM | ? | $c_3$ |  |
| SCD | ? | $c_4$ |  |
| All-cause death | ? | $c_5$ |  |
|      |      |      |     |

: Model parameter values.

@@

This gives us starting state populations for HCM with ICD and HCM without ICD which will be different for each risk decision rule.
Further, the transition probabilities from these states will differ because of the case mixes.

@@

We have used a year step size but if eg cost are accrued at different intervals then this can be adapted.
The time horizon is set at 12 years but can be changed.

@@

Using WinBUGS called from R, each new data set is used to generate posterior samples of transition probabilities.
Denote $x$ as the observed number of transitions, $p$ the probability of a transition and $n$ as the total number of transitions from a given state. The hyperparameters $\alpha$ characterise the prior knowledge on $p$.
Superscripts indicate the decision rule used.

>>>>

---
title: "HCM SCD risk algorithm cost-effectiveness analysis"
author: "Nathan Green"
date: "09/11/2020"
output: pdf_document
editor_options: 
  markdown: 
    wrap: sentence
bibliography: bibliography.bib
---

@@

Sudden cardiac death (SCD) risk score algorithm details see [@OMahony2014].

There have been previous cost-effectiveness analyses [@Magnusson2020; @Boriani2014; @Bryant2007; @Caro2007; @Cowie2009; @Garcia-Perez2015; @Mealing2016; @Sanders2005; @Smith2013]

@@

+----------------------------------+-----------------+--------+------------------+
| Description                      | Value$^*$       | Range  | Source           |
+==================================+=================+========+==================+
| *Health*                         |                 |        |                  |
+----------------------------------+-----------------+--------+------------------+
| Manage with ICD                  | 0.637 QALY/year |        | @Noyes2007       |
+----------------------------------+-----------------+--------+------------------+
| Implantation procedure utility   | -0.016          |        | @Smith2013       |
+----------------------------------+-----------------+--------+------------------+
| Shock utility                    | -0.5            |        |                  |
+----------------------------------+-----------------+--------+------------------+
| HCM without ICD                  | 1 QALY/year     |        |                  |
+----------------------------------+-----------------+--------+------------------+
| Death                            | 0 QALY/year     |        |                  |
+----------------------------------+-----------------+--------+------------------+
|                                  |                 |        |                  |
+----------------------------------+-----------------+--------+------------------+
| *Cost*                           |                 |        |                  |
+----------------------------------+-----------------+--------+------------------+
| ICD appointment                  | £10             |        |                  |
+----------------------------------+-----------------+--------+------------------+
| Perform risk score               | £20?            |        |                  |
+----------------------------------+-----------------+--------+------------------+
| Implant ICD                      | £4,666          |        | EY02B Tariffs    |
+----------------------------------+-----------------+--------+------------------+
| Implant replacement              | £45,000         |        |                  |
+----------------------------------+-----------------+--------+------------------+
| Implant complication             | £28,839         |        | @Smith2013       |
+----------------------------------+-----------------+--------+------------------+
| Non-fatal shock                  | £22,880         |        | UK Stroke Assoc. |
+----------------------------------+-----------------+--------+------------------+
| HCM without ICD                  | 0               |        |                  |
+----------------------------------+-----------------+--------+------------------+
| SCD                              | 0               |        |                  |
+----------------------------------+-----------------+--------+------------------+
| All-cause death                  | 0               |        |                  |
+----------------------------------+-----------------+--------+------------------+
|                                  |                 |        |                  |
+----------------------------------+-----------------+--------+------------------+
| *Probabilities*                  |                 |        |                  |
+----------------------------------+-----------------+--------+------------------+
| Initial implant complication     | 0.047           |        | @Smith2013       |
+----------------------------------+-----------------+--------+------------------+
| Replacement implant complication | 0.032           |        | @Smith2013       |
+----------------------------------+-----------------+--------+------------------+
|                                  |                 |        |                  |
+----------------------------------+-----------------+--------+------------------+
| Time horizon                     | 12 years        |        |                  |
+----------------------------------+-----------------+--------+------------------+
| Implant replacement              | 10 years        |        |                  |
+----------------------------------+-----------------+--------+------------------+
| Annual number of appointments    | 2               |        |                  |
+----------------------------------+-----------------+--------+------------------+

: Model parameter values. $^*$either one-off/on state entry or recurring.

Table 2 gives the starting state populations for non-zero states.

+-----------------+---------+------------+
| Risk rule       | State   | Population |
+=================+=========+============+
| Observed        | HCM ICD | 559        |
+-----------------+---------+------------+
| \-              | HCM     | 3113       |
+-----------------+---------+------------+
| Score \> 4%     | HCM ICD | 2561       |
+-----------------+---------+------------+
| \-              | HCM     | 1111       |
+-----------------+---------+------------+
| Score \> 6%     | HCM ICD | 542        |
+-----------------+---------+------------+
| \-              | HCM     | 3130       |
+-----------------+---------+------------+
| Risk factor \>0 | HCM ICD | 1785       |
+-----------------+---------+------------+
| \-              | HCM     | 1887       |
+-----------------+---------+------------+
| Risk factor \>1 | HCM ICD | 481        |
+-----------------+---------+------------+
| \-              | HCM     | 3191       |
+-----------------+---------+------------+
| Risk factor \>2 | HCM ICD | 78         |
+-----------------+---------+------------+
| \-              | HCM     | 3594       |
+-----------------+---------+------------+

@@

### Markov model

Th patient data give us starting state populations for HCM with ICD and HCM without ICD which will be different for each risk decision rule.
Further, the transition probabilities from these states will differ because of the case mixes.

@@

We have used a year step size but if e.g. cost are accrued at different intervals then this can be adapted.
The time horizon is set at 12 years from time of implant.

### State health and cost equations

We assume that an ICD patient has 2 annual appointments.
All shocks are treated the same in terms of costs and health impact.
Implantation can have complications.
The cost of an implant complication is taken as a weighted sum of infection and dislodgement cost with values from [@Smith2013].

-   Annual health: $$
    e^0_{s=1}(t) = e^1_{s=1}(t) = u\_hcm + u\_icd
    $$

$$
e^0_{s=2}(t) = e^1_{s=2}(t) = u\_shock
$$

$$
e^0_{s}(t) = e^1_{s}(t) = 0, \;\; s = 3,4,5,6
$$

-   Initial costs: $$
    c^0_{s=1}(t = 0) = c\_icd + p\_compl \times c\_compl
    $$

$$
c^1_{s=1}(t = 0) = c\_icd + c\_rscore + p\_compl \times c\_compl
$$

-   Annual cost: $$
    c^0_{s=1}(t) = c^1_{s=1}(t) = 2 c\_icd\_appt, \;\; t \neq t\_repl
    $$

$$
c^0_{s=2}(t) = c^1_{s=2}(t) = c\_nfatal
$$

-   Implant replacement cost $$
    c^0_{s=1}(t = t\_repl) = c^1_{s=1}(t = t\_repl) = 2 c\_icd\_appt + c\_icd\_repl + p\_compl \times c\_compl
    $$

@@

Using WinBUGS [@Zhang2014] called from R, each new data set is used to generate posterior samples of transition probabilities.
Denote $x$ as the observed number of transitions, $p$ the probability of a transition and $n$ as the total number of transitions from a given state.
The hyperparameters $\alpha$ characterise the prior knowledge on $p$.
Superscripts indicate the decision rule used.
